---
title: "p8105_hw3_yh3964"
author: "Yirou Hu"
date: "2025-10-08"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and datasets
```{r}
library(patchwork)
library(tidyverse)
library(dplyr)
library(p8105.datasets)
data("instacart") 
```

Set the size of plots

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```

Change legend position to bottom and remove the frame

```{r}
theme_set(theme_minimal() + theme(legend.position = "bottom")) 
```

We change continuous and discrete variables with "viridis" for color and fill

```{r}
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each)

## How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisles_number_df = 
  instacart |>
  group_by(aisle) |>
  summarize(count_aisle = n()) |>
  mutate(aisle_rank = rank(desc(count_aisle))) |>
  arrange(aisle_rank)
```

Dataset description:

The whole dataset provides detailed information on each product that customers ordered, including the product name, department, and aisle, as well as the information about the time they make this order, such as order_dow and order_hour_of_day. Additionally, it includes informaiton about whether the product was reordered, the order number, and the number of days since the prior order. The product id, aisle id and the order they belonged to are also provided.

There are `r length(unique(instacart$aisle))` aisles and `r aisles_number_df$aisle[which(aisles_number_df$aisle_rank ==1)]` is the aisle with the most items ordered from.

## Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r, fig.width = 20, fig.height = 25}
instacart |>
  group_by(aisle) |>
  summarize(count_aisle = n()) |>
  arrange(count_aisle) |>
  filter(count_aisle > 10000) |>
  ggplot(aes(x = count_aisle, y = reorder(aisle, -count_aisle))) +
  geom_bar(stat = "identity", fill = "blue", alpha = 0.5) +
  labs(                                             
    title = "Number of Items Ordered in Each Aisle",
    x = "Number of Items Ordered",
    y = "Aisle"
  ) 
```

Plot description:

This bar chart shows the number of items ordered from each aisle. It's clear that "fresh vegetables" and "fresh fruits" are the most popular, with the highest order counts. Other top sellers include "packaged vegetables and fruits," "yogurt," and "packaged cheese." Less ordered items come from aisles like "butter" and "oils vinegars," yet they still exceed 10,000 orders. This chart helps identify which products are in highest demand.

## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
popular_items_df = 
  instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle, product_name) |>
  summarise(number_of_time = n()) |>
  mutate(rank_of_time = rank(desc(number_of_time))) |>
  arrange(rank_of_time) |>
  filter(rank_of_time == c(1,2,3)) |>
  arrange(desc(number_of_time)) 

popular_items_df |>
  knitr::kable(col.names = c("Aisle", "Product Name", "Number of times ordered", "Rank"))
```

## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarise(mean_order_hour_of_day = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_order_hour_of_day
  ) |>
  knitr::kable() 
```

The table shows the average ordering hours for "Pink Lady Apples" and "Coffee Ice Cream" by day of the week. "Coffee Ice Cream" orders peak mid-week, especially on Wednesdays around 15 hours. "Pink Lady Apples" are most ordered on Thursdays around 14 hours. 










